\section{Notation and basic definitions}

Sets in this paper are denoted by Euler script letters,
for example $\mathcal{I}$.
Small bold letters like $\pmb{u}$ are for vectors.
Capital letters mean matrices, e. g. $A$.
Upper indices of matrices mean columns, lower are for rows.
Row is a vector as well as a column,
so they are denoted by the same letter as corresponding matrix,
but bold lowercase.
For example, row $0$ of matrix $A$ is $\pmb{a}_0$,
column $x$ of matrix $\Lambda$ is $\pmb{\lambda}^x$.
Vectorized matrix consists of columns of the matrix stacked together
one by one and will be denoted by bold capital letter
\begin{equation*}
  A = \begin{bmatrix}
    a^1_1  & a^1_2  & \dots & a^1_m  \\
    a^2_1  & a^2_2  & \dots & a^2_m  \\
    \vdots & \vdots & \dots & \vdots \\
    a^n_1  & a^n_2  & \dots & a^n_m  \\
  \end{bmatrix}
  \mapsto
  \begin{bmatrix}
    a^1_1     \\
    a^1_2     \\
    \vdots    \\
    a^1_m     \\
    a^2_1     \\
    a^2_2     \\
    \vdots    \\
    a^n_{m-1} \\
    a^n_m     \\
  \end{bmatrix}
  = \pmb{A}
\end{equation*}

3D model consists of vertices.
Each vertex has its own tuple of properties.
It's convenient to use vertex identifier as an index for its properties instead of vice versa.
The set of vertices will be denoted $V$ and its cardinality is $m$.
Particular vertices will be $v$, $u$ and $w$~---~in this document
it's assumed that these letters in formulas
mean any element from the set $V$ unless otherwise specified.

We need three sets of indices for space axes:
\begin{itemize}
  \item $\mathcal{I} = \left\{ x, y, z \right\}$ for ordinary 3D space;
  \item $\mathcal{I}^h = \left\{ 0, x, y, z \right\}$ for homogeneous coordinates;
  \item $\mathcal{I}^p = \left\{ x, y \right\}$ for plane.
\end{itemize}

Projection of vertex $v$ on $i$ axis is $v_i$ and it's a scalar.
Projection of vector $\pmb{v}$ on vector $\pmb{u}$ is a vector
\begin{equation*}
  \pmb{v} \parallel \pmb{u}
  = \frac{\left( \pmb{v}, \pmb{u} \right)}{\left\| \pmb{u} \right\|^2}
    \cdot \pmb{u}.
\end{equation*}
